---
output: 
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    number_sections: yes
geometry: margin=2.54cm
title: "Is Physical Activity Associated with C-protein Level in US Population with Different Framingham Scores"
subtitle: "https://github.com/YinsuW/Wang1Asfaw2Qian3_ENV872_EDA_FinalProject.git"
author: "Yinsu Wang, Yared Asfaw, April Qian"
fontsize: 12pt
mainfont: Times New Roman

---

\newpage
\tableofcontents 
\newpage
\listoftables 
\newpage
\listoffigures 
\newpage

```{r setup, include=FALSE}
# Set your working directory
setwd("/Users/wangyinsu/Desktop/final\ project/data")
# Load your packages
library(tidyverse)
library(ggplot2)
library(foreign)
library(data.table)
library(naniar)
library(tidyr)
library(naniar)
library(caret)
library(mgcv)
# Set your ggplot theme
mytheme <- theme_classic() +
  theme(axis.text = element_text(color = "black"), 
        legend.position = "top")
theme_set(mytheme)
# Load your datasets
rm(list=ls())
temp = list.files(path = "/Users/wangyinsu/Desktop/final\ project/data", pattern="*.XPT")
myfiles = lapply(temp, read.xport)
PA.raw<-rbindlist(myfiles, use.names=TRUE, fill=TRUE)

srvyin <- paste("1999_2000.dat") 
srvyout <- "1999"
dsn <- read_fwf(file=srvyin,
                col_types = "ciiiiiiiddii",
                fwf_cols(publicid = c(1,14),
                         eligstat = c(15,15),
                         mortstat = c(16,16),
                         ucod_leading = c(17,19),
                         diabetes = c(20,20),
                         hyperten = c(21,21),
                         dodqtr = c(22,22),
                         dodyear = c(23,26),
                         wgt_new = c(27,34),
                         sa_wgt_new = c(35,42),
                         permth_int = c(43,45),
                         permth_exm = c(46,48)
                ),
                na = "."
)
dsn$seqn <- substr(dsn$publicid,1,5)
dsn <- select(dsn, -publicid)
dsn <- select(dsn, -dodqtr)
dsn <- select(dsn, -dodyear)
dsn <- select(dsn, -wgt_new)
dsn <- select(dsn, -sa_wgt_new)
str(dsn)
table(dsn$eligstat)
table(dsn$mortstat, useNA="ifany")
table(dsn$ucod_leading, useNA="ifany")
table(dsn$diabetes, useNA="ifany")
table(dsn$hyperten, useNA="ifany")
assign(paste0(srvyout), dsn)

srvyin2 <- paste("2001_2002.dat")  
srvyout2 <- "2001"
dsn2 <- read_fwf(file=srvyin2,
                col_types = "ciiiiiiiddii",
                fwf_cols(publicid = c(1,14),
                         eligstat = c(15,15),
                         mortstat = c(16,16),
                         ucod_leading = c(17,19),
                         diabetes = c(20,20),
                         hyperten = c(21,21),
                         dodqtr = c(22,22),
                         dodyear = c(23,26),
                         wgt_new = c(27,34),
                         sa_wgt_new = c(35,42),
                         permth_int = c(43,45),
                         permth_exm = c(46,48)
                ),
                na = "."
)
dsn2$seqn <- substr(dsn2$publicid,1,5)
dsn2 <- select(dsn2, -publicid)
dsn2 <- select(dsn2, -dodqtr)
dsn2 <- select(dsn2, -dodyear)
dsn2 <- select(dsn2, -wgt_new)
dsn2 <- select(dsn2, -sa_wgt_new)
str(dsn2)
table(dsn2$eligstat)
table(dsn2$mortstat, useNA="ifany")
table(dsn2$ucod_leading, useNA="ifany")
table(dsn2$diabetes, useNA="ifany")
table(dsn2$hyperten, useNA="ifany")
assign(paste0(srvyout2), dsn2)

srvyin3 <- paste("2003_2004.dat")  
srvyout3 <- "2003"
dsn3 <- read_fwf(file=srvyin3,
                col_types = "ciiiiiiiddii",
                fwf_cols(publicid = c(1,14),
                         eligstat = c(15,15),
                         mortstat = c(16,16),
                         ucod_leading = c(17,19),
                         diabetes = c(20,20),
                         hyperten = c(21,21),
                         dodqtr = c(22,22),
                         dodyear = c(23,26),
                         wgt_new = c(27,34),
                         sa_wgt_new = c(35,42),
                         permth_int = c(43,45),
                         permth_exm = c(46,48)
                ),
                na = "."
)
dsn3$seqn <- substr(dsn3$publicid,1,5)
dsn3 <- select(dsn3, -publicid)
dsn3 <- select(dsn3, -dodqtr)
dsn3 <- select(dsn3, -dodyear)
dsn3 <- select(dsn3, -wgt_new)
dsn3 <- select(dsn3, -sa_wgt_new)
str(dsn3)
table(dsn3$eligstat)
table(dsn3$mortstat, useNA="ifany")
table(dsn3$ucod_leading, useNA="ifany")
table(dsn3$diabetes, useNA="ifany")
table(dsn3$hyperten, useNA="ifany")
assign(paste0(srvyout3), dsn3)

srvyin4 <- paste("2005_2006.dat")  
srvyout4 <- "2005"
dsn4 <- read_fwf(file=srvyin4,
                col_types = "ciiiiiiiddii",
                fwf_cols(publicid = c(1,14),
                         eligstat = c(15,15),
                         mortstat = c(16,16),
                         ucod_leading = c(17,19),
                         diabetes = c(20,20),
                         hyperten = c(21,21),
                         dodqtr = c(22,22),
                         dodyear = c(23,26),
                         wgt_new = c(27,34),
                         sa_wgt_new = c(35,42),
                         permth_int = c(43,45),
                         permth_exm = c(46,48)
                ),
                na = "."
)
dsn4$seqn <- substr(dsn4$publicid,1,5)
dsn4 <- select(dsn4, -publicid)
dsn4 <- select(dsn4, -dodqtr)
dsn4 <- select(dsn4, -dodyear)
dsn4 <- select(dsn4, -wgt_new)
dsn4 <- select(dsn4, -sa_wgt_new)
str(dsn4)
table(dsn4$eligstat)
table(dsn4$mortstat, useNA="ifany")
table(dsn4$ucod_leading, useNA="ifany")
table(dsn4$diabetes, useNA="ifany")
table(dsn4$hyperten, useNA="ifany")
assign(paste0(srvyout4), dsn4)

srvyin5 <- paste("2007_2008.dat")  
srvyout5 <- "2007"
dsn5 <- read_fwf(file=srvyin5,
                col_types = "ciiiiiiiddii",
                fwf_cols(publicid = c(1,14),
                         eligstat = c(15,15),
                         mortstat = c(16,16),
                         ucod_leading = c(17,19),
                         diabetes = c(20,20),
                         hyperten = c(21,21),
                         dodqtr = c(22,22),
                         dodyear = c(23,26),
                         wgt_new = c(27,34),
                         sa_wgt_new = c(35,42),
                         permth_int = c(43,45),
                         permth_exm = c(46,48)
                ),
                na = "."
)
dsn5$seqn <- substr(dsn5$publicid,1,5)
dsn5 <- select(dsn5, -publicid)
dsn5 <- select(dsn5, -dodqtr)
dsn5 <- select(dsn5, -dodyear)
dsn5 <- select(dsn5, -wgt_new)
dsn5 <- select(dsn5, -sa_wgt_new)
str(dsn5)
table(dsn5$eligstat)
table(dsn5$mortstat, useNA="ifany")
table(dsn5$ucod_leading, useNA="ifany")
table(dsn5$diabetes, useNA="ifany")
table(dsn5$hyperten, useNA="ifany")
assign(paste0(srvyout5), dsn5)

srvyin6 <- paste("2009_2010.dat")  
srvyout6 <- "2009"
dsn6 <- read_fwf(file=srvyin6,
                col_types = "ciiiiiiiddii",
                fwf_cols(publicid = c(1,14),
                         eligstat = c(15,15),
                         mortstat = c(16,16),
                         ucod_leading = c(17,19),
                         diabetes = c(20,20),
                         hyperten = c(21,21),
                         dodqtr = c(22,22),
                         dodyear = c(23,26),
                         wgt_new = c(27,34),
                         sa_wgt_new = c(35,42),
                         permth_int = c(43,45),
                         permth_exm = c(46,48)
                ),
                na = "."
)
dsn6$seqn <- substr(dsn6$publicid,1,5)
dsn6 <- select(dsn6, -publicid)
dsn6 <- select(dsn6, -dodqtr)
dsn6 <- select(dsn6, -dodyear)
dsn6 <- select(dsn6, -wgt_new)
dsn6 <- select(dsn6, -sa_wgt_new)
str(dsn6)
table(dsn6$eligstat)
table(dsn6$mortstat, useNA="ifany")
table(dsn6$ucod_leading, useNA="ifany")
table(dsn6$diabetes, useNA="ifany")
table(dsn6$hyperten, useNA="ifany")
assign(paste0(srvyout6), dsn6)

CVD <-rbind(dsn, dsn2,dsn3,dsn4,dsn5,dsn6)
CVD<-rename(CVD, SEQN = seqn)

nhanes<-merge(CVD,PA.raw)
```


# Rationale and Research Questions
In this final project, we aims to explore the relationship between physical activity (PA) and C-protein level in population with different framingham scores using NHANES datasets from 1999 to 2010.

##Question 1:
Is physical activity associated with the level of C-protein? If so, what is the exact relationship? Is it dose-response? 

##Question 2: 
Are different levels of framingham scores the effect modifiers for this relationship between physical activity and C-protein? How do they modify the relationship?


\newpage

# Dataset Information
NHANES is National Health and Nitrition Examination Survey used to assess health and nutrition status of adults and children in the United States. The examination component consists of medical, dental, and physiological measurements, as well as laboratory tests administered by highly trained medical personnel. In this project, we will use this NHANES dataset to investigate into the relationship between PA and C-protein level in since we hypothese that PA is related to C-protein level in Patients Having different levels of framingham scores.

##Methods
*1.	*Inclusion and Exclusion Criteria:
+ We include participants whose ages are equal or above 20 years with a diagnosis (I00-I09, I11, I13, I20-I51) of disease of heart using NHANES datasets from 1999 to 2010. 
+ We exclude participants whose cholesterol (HDL, TC), blood pressure, diabetes, PA status, race, education level, family PIR are missing. 
*2. *Statistical Analysis
+ We used R software to examine the association between physical activity and c-protein level in the study population. Multiple logistic regression was applied to study the association between the two. We considered physical activity intensity of “walk or bicycle” as the control group, while moderate- and high-intensity work as the cases. We identified confounders as cholesterol (HDL, TC), blood pressure, diabetes, race, education level, family PIR and these variables was controlled in the analysis. 

\newpage

# Exploratory Analysis 

```{r echo=FALSE, message=FALSE, warning=FALSE}
##Data wrangling

nhanes1<-subset(nhanes,select = c("SEQN",'RIDAGEYR','LBDHDDSI','LBDTCSI','BPQ040A','SMQ040','DIQ010','PAQ605','PAQ610','PAD615','PAQ620','PAQ625','PAD630','PAQ635','PAQ640','PAD645','RIDRETH1','DMDEDUC2','INDFMPIR','BMXBMI','LBXCRP','ucod_leading',"BPXSY1","BPXSY2","BPXSY3","RIAGENDR"))
nhanes1<-setDT(nhanes1)[, lapply(.SD, na.omit), by = SEQN]
nhanes1<-unique(nhanes1)

nhanes1<-nhanes1 %>% replace_with_na(replace = list(PAQ605 = 9))%>% 
  replace_with_na(replace = list(PAQ610 = 99))%>% 
  replace_with_na(replace = list(PAD615 = 9999))%>% 
  replace_with_na(replace = list(PAQ620 = 9))%>% 
  replace_with_na(replace = list(PAQ625 = 99))%>% 
  replace_with_na(replace = list(PAD630 = 9999))%>% 
  replace_with_na(replace = list(PAQ635 = 9))%>% 
  replace_with_na(replace = list(PAQ640 = 99))%>% 
  replace_with_na(replace = list(PAD645 = 9999))
nhanes1<-nhanes1 %>% replace_with_na(replace = list(PAQ605 = 7))%>% 
  replace_with_na(replace = list(PAQ610 = 77))%>% 
  replace_with_na(replace = list(PAD615 = 7777))%>% 
  replace_with_na(replace = list(PAQ620 = 7))%>% 
  replace_with_na(replace = list(PAQ625 = 77))%>% 
  replace_with_na(replace = list(PAD630 = 7777))%>% 
  replace_with_na(replace = list(PAQ635 = 7))%>% 
  replace_with_na(replace = list(PAQ640 = 77))%>% 
  replace_with_na(replace = list(PAD645 = 7777))
nhanes1<-nhanes1[!is.na(PAQ605)|!is.na(PAQ620)|is.na(PAQ635)]
nhanes1<-nhanes1%>%mutate(vigorous = (nhanes1$PAD615)*(nhanes1$PAQ610)*8)%>%
  mutate(moderate_walk = (nhanes1$PAD645)*(nhanes1$PAQ640)*4)%>%
  mutate(moderate_activity = (nhanes1$PAQ625)*(nhanes1$PAD630)*4)
nhanes1<-subset(nhanes1,is.na(moderate_activity)==FALSE | is.na(moderate_walk)==FALSE | is.na(vigorous)==FALSE)

nhanes1 <- nhanes1 %>% mutate(SBP = 
                            case_when(BPXSY1>0&is.na(BPXSY2)&is.na(BPXSY3) ~ BPXSY1,
                                      BPXSY1>0&BPXSY2>0&is.na(BPXSY3) ~ BPXSY2,
                                      BPXSY1>0&BPXSY2>0&BPXSY3>0 ~ (BPXSY2+BPXSY3)/2))

nhanes2<-nhanes1[complete.cases(nhanes1[,c('RIDAGEYR','LBDHDDSI','LBDTCSI','SMQ040','DIQ010','SBP')]),]
nhanes2=filter(filter(filter(filter(filter(filter(nhanes2,SMQ040 != 7),SMQ040 != 9),DIQ010 != 7),DIQ010 != 9),BPQ040A !=7),BPQ040A!=9)
nhanes2=filter(nhanes2,RIDAGEYR>=20)

nhanes21<-subset(nhanes2,select = c("SEQN",'RIDAGEYR','LBDHDDSI','LBDTCSI','BPQ040A','SMQ040','DIQ010','PAQ605','PAQ610','PAD615','PAQ620','PAQ625','PAD630','PAQ635','PAQ640','PAD645','RIDRETH1','DMDEDUC2','INDFMPIR','BMXBMI','LBXCRP','ucod_leading',"BPXSY1","BPXSY2","BPXSY3","RIAGENDR","vigorous","SBP"))
nhanes21<-rename(nhanes21, PA = vigorous)

nhanes22<-subset(nhanes2,select = c("SEQN",'RIDAGEYR','LBDHDDSI','LBDTCSI','BPQ040A','SMQ040','DIQ010','PAQ605','PAQ610','PAD615','PAQ620','PAQ625','PAD630','PAQ635','PAQ640','PAD645','RIDRETH1','DMDEDUC2','INDFMPIR','BMXBMI','LBXCRP','ucod_leading',"BPXSY1","BPXSY2","BPXSY3","RIAGENDR","moderate_activity","SBP"))
nhanes22<-rename(nhanes22, PA = moderate_activity)
nhanes23<-subset(nhanes2,select = c("SEQN",'RIDAGEYR','LBDHDDSI','LBDTCSI','BPQ040A','SMQ040','DIQ010','PAQ605','PAQ610','PAD615','PAQ620','PAQ625','PAD630','PAQ635','PAQ640','PAD645','RIDRETH1','DMDEDUC2','INDFMPIR','BMXBMI','LBXCRP','ucod_leading',"BPXSY1","BPXSY2","BPXSY3","RIAGENDR","moderate_walk","SBP"))
nhanes23<-rename(nhanes23, PA = moderate_walk)
nhanes3<-rbind(nhanes21,nhanes22,nhanes23)
nhanes3<-nhanes3[!is.na(nhanes3$PA)]

# framingham score
if (nhanes3$RIAGENDR==1){nhanes3<-nhanes3%>%mutate(score_Age = 
  case_when  
           ((nhanes3$RIDAGEYR >=30) & (nhanes3$RIDAGEYR<=34) ~ 0,
            (nhanes3$RIDAGEYR >=35) & (nhanes3$RIDAGEYR <=39) ~2,
            (RIDAGEYR >=40) & (RIDAGEYR <=44) ~5,
            (RIDAGEYR >=45) & (RIDAGEYR <=49) ~7,
            (RIDAGEYR >=50) & (RIDAGEYR <=54) ~8,
            (RIDAGEYR >=55) & (RIDAGEYR <=59) ~10,
            (RIDAGEYR >=60) & (RIDAGEYR <=64) ~11,
            (RIDAGEYR >=65) & (RIDAGEYR <=69) ~12,
            (RIDAGEYR >=70) & (RIDAGEYR <=74) ~14,
            (RIDAGEYR >=75) ~15))}else{nhanes3%>%mutate(score_Age = 
  case_when  
  ((RIDAGEYR >=30) & (RIDAGEYR <=34) ~ 0,
            (RIDAGEYR >=35) & (RIDAGEYR <=39) ~2,
            (RIDAGEYR >=40) & (RIDAGEYR <=44) ~4,
            (RIDAGEYR >=45) & (RIDAGEYR <=49) ~5,
            (RIDAGEYR >=50) & (RIDAGEYR <=54) ~7,
            (RIDAGEYR >=55) & (RIDAGEYR <=59) ~8,
            (RIDAGEYR >=60) & (RIDAGEYR <=64) ~9,
            (RIDAGEYR >=65) & (RIDAGEYR <=69) ~10,
            (RIDAGEYR >=70) & (RIDAGEYR <=74) ~11,
            (RIDAGEYR >=75) ~12))}

if (nhanes3$RIAGENDR==1){nhanes3<-nhanes3%>%mutate(score_HDL = 
    case_when((nhanes3$LBDHDDSI>1.6) ~ -2,
              (nhanes3$LBDHDDSI>=1.3) & (nhanes3$LBDHDDSI <=1.6) ~ -1,
              (nhanes3$LBDHDDSI>=1.2) & (nhanes3$LBDHDDSI <=1.29) ~ 0,
              (nhanes3$LBDHDDSI>=0.9) & (nhanes3$LBDHDDSI <=1.19) ~ 1,
              (nhanes3$LBDHDDSI<0.9) ~ 2))}else{nhanes3<-nhanes3%>%mutate(score_HDL = 
    case_when((nhanes3$LBDHDDSI>1.6) ~ -2,
              (nhanes3$LBDHDDSI>=1.3) & (nhanes3$LBDHDDSI <=1.6) ~ -1,
              (nhanes3$LBDHDDSI>=1.2) & (nhanes3$LBDHDDSI <=1.29) ~ 0,
              (nhanes3$LBDHDDSI>=0.9) & (nhanes3$LBDHDDSI <=1.19) ~ 1,
              (nhanes3$LBDHDDSI<0.9) ~ 2))}
if (nhanes3$RIAGENDR==1){nhanes3<-nhanes3%>%mutate(score_TC = 
    case_when((nhanes3$LBDTCSI>7.2) ~ 4,
              (nhanes3$LBDTCSI>=6.2) & (nhanes3$LBDTCSI <=7.2) ~ 3,
              (nhanes3$LBDTCSI>=5.2) & (nhanes3$LBDTCSI <=6.19) ~ 2,
              (nhanes3$LBDTCSI>=4.1) & (nhanes3$LBDTCSI <=5.19) ~ 1,
              (nhanes3$LBDTCSI<4.1) ~ 0))}else{nhanes3<-nhanes3%>%mutate(score_TC = 
    case_when((nhanes3$LBDTCSI>7.2) ~ 5,
              (nhanes3$LBDTCSI>=6.2) & (nhanes3$LBDTCSI <=7.2) ~ 4,
              (nhanes3$LBDTCSI>=5.2) & (nhanes3$LBDTCSI <=6.19) ~ 3,
              (nhanes3$LBDTCSI>=4.1) & (nhanes3$LBDTCSI <=5.19) ~ 1,
              (nhanes3$LBDTCSI<4.1) ~ 0))}

if (nhanes3$RIAGENDR==1){nhanes3<-nhanes3%>%mutate(score_BP = 
   if(nhanes3$BPQ040A==1) {case_when((nhanes3$SBP>160) ~ 5,
              (nhanes3$SBP>=150) & (nhanes3$SBP <=159) ~ 4,
              (nhanes3$SBP>=140) & (nhanes3$SBP <=149) ~ 4,
              (nhanes3$SBP>=130) & (nhanes3$SBP <=139) ~ 3,
              (nhanes3$SBP>=120) & (nhanes3$SBP <=129) ~ 2,
              (nhanes3$SBP<120) ~ 0)}else
   if(nhanes3$BPQ040A==2) {case_when((nhanes3$SBP>160) ~ 3,
              (nhanes3$SBP>=150) & (nhanes3$SBP <=159) ~ 2,
              (nhanes3$SBP>=140) & (nhanes3$SBP <=149) ~ 2,
              (nhanes3$SBP>=130) & (nhanes3$SBP <=139) ~ 1,
              (nhanes3$SBP>=120) & (nhanes3$SBP <=129) ~ 0,
              (nhanes3$SBP<120) ~ -2)})}else{nhanes3<-nhanes3%>%mutate(score_BP = 
   if(nhanes3$BPQ040A==1) {case_when((nhanes3$SBP>160) ~ 7,
              (nhanes3$SBP>=150) & (nhanes3$SBP <=159) ~ 6,
              (nhanes3$SBP>=140) & (nhanes3$SBP <=149) ~ 5,
              (nhanes3$SBP>=130) & (nhanes3$SBP <=139) ~ 3,
              (nhanes3$SBP>=120) & (nhanes3$SBP <=129) ~ 2,
              (nhanes3$SBP<120) ~ -1)}else
   if(nhanes3$BPQ040A==2) {case_when((nhanes3$SBP>160) ~ 5,
              (nhanes3$SBP>=150) & (nhanes3$SBP <=159) ~ 4,
              (nhanes3$SBP>=140) & (nhanes3$SBP <=149) ~ 2,
              (nhanes3$SBP>=130) & (nhanes3$SBP <=139) ~ 1,
              (nhanes3$SBP>=120) & (nhanes3$SBP <=129) ~ 0,
              (nhanes3$SBP<120) ~ -3)})}

   if(nhanes3$RIAGENDR==2){nhanes3<-nhanes3%>%mutate(score_Smoke =
                               case_when((nhanes3$SMQ040==1) ~ 3,
                                         (nhanes3$SMQ040==2) ~ 3,
                                         (nhanes3$SMQ040==3) ~ 0))}else
                                         {nhanes3<-nhanes3%>%mutate(score_Smoke =
                               case_when((nhanes3$SMQ040==1) ~ 4,
                                         (nhanes3$SMQ040==2) ~ 4,
                                         (nhanes3$SMQ040==3) ~ 0))}
   if(nhanes3$RIAGENDR==2){nhanes3<-nhanes3%>%mutate(score_Diabetes =
                               case_when((nhanes3$SMQ040==1) ~ 4,
                                         (nhanes3$SMQ040==2) ~ 4,
                                         (nhanes3$SMQ040==3) ~ 0))}else
                                         {nhanes3<-nhanes3%>%mutate(score_Diabetes =
                               case_when((nhanes3$SMQ040==1) ~ 3,
                                         (nhanes3$SMQ040==2) ~ 3,
                                         (nhanes3$SMQ040==3) ~ 0))}
nhanes3<-nhanes3%>%mutate(score=score_Age+score_BP+score_HDL+score_TC+score_Smoke+score_Diabetes)

   if(nhanes3$RIAGENDR==1){nhanes3<-nhanes3%>%mutate(risk =
                               case_when((nhanes3$score <= -3) ~ 1,
                                         (nhanes3$score==-2) ~ 1.1,
                                         (nhanes3$score==-1) ~ 1.4,
                                         (nhanes3$score==0) ~ 1.6,
                                         (nhanes3$score==1) ~ 1.9,
                                         (nhanes3$score==2) ~ 2.3,
                                         (nhanes3$score==3) ~ 2.8,
                                         (nhanes3$score==4) ~ 3.3,
                                         (nhanes3$score==5) ~ 3.9,
                                         (nhanes3$score==6) ~ 4.7,
                                         (nhanes3$score==7) ~ 5.6,
                                         (nhanes3$score==8) ~ 6.7,
                                         (nhanes3$score==9) ~ 7.9,
                                         (nhanes3$score==10) ~ 9.4,
                                         (nhanes3$score==11) ~ 11.2,
                                         (nhanes3$score==12) ~ 13.3,
                                         (nhanes3$score==13) ~ 15.6,
                                         (nhanes3$score==14) ~ 18.4,
                                         (nhanes3$score==15) ~ 21.6,
                                         (nhanes3$score==16) ~ 25.3,
                                         (nhanes3$score==17) ~ 29.4,
                                         (nhanes3$score==18) ~ 30,
                                         (nhanes3$score==19) ~ 30,
                                         (nhanes3$score==20) ~ 30,
                                         (nhanes3$score>=21) ~ 30,))}else
                                         {nhanes3<-nhanes3%>%mutate(risk =
                               case_when((nhanes3$score <= -3) ~ 1,
                                         (nhanes3$score==-2) ~ 1,
                                         (nhanes3$score==-1) ~ 1,
                                         (nhanes3$score==0) ~ 1.2,
                                         (nhanes3$score==1) ~ 1.5,
                                         (nhanes3$score==2) ~ 1.7,
                                         (nhanes3$score==3) ~ 2.0,
                                         (nhanes3$score==4) ~ 2.4,
                                         (nhanes3$score==5) ~ 2.8,
                                         (nhanes3$score==6) ~ 3.3,
                                         (nhanes3$score==7) ~ 3.9,
                                         (nhanes3$score==8) ~ 4.5,
                                         (nhanes3$score==9) ~ 5.3,
                                         (nhanes3$score==10) ~ 6.3,
                                         (nhanes3$score==11) ~ 7.3,
                                         (nhanes3$score==12) ~ 8.6,
                                         (nhanes3$score==13) ~ 10.0,
                                         (nhanes3$score==14) ~ 11.7,
                                         (nhanes3$score==15) ~ 13.7,
                                         (nhanes3$score==16) ~ 15.9,
                                         (nhanes3$score==17) ~ 18.51,
                                         (nhanes3$score==18) ~ 21.5,
                                         (nhanes3$score==19) ~ 24.8,
                                         (nhanes3$score==20) ~ 27.5,
                                         (nhanes3$score>=21) ~ 30,))}

#risk_level definition
if (nhanes3$RIAGENDR==1){nhanes3<-nhanes3%>%mutate(risk_level = 
  case_when  
           ((risk <=10) ~ "low",
            (risk >10) & (risk <=14) ~"intermediate",
            (risk >14) ~"high"))}else{nhanes3%>%mutate(risk_level = 
  case_when  
  ((risk <=12) ~ "low",
            (risk >12) & (risk <=17) ~"intermediate",
            (risk >17) ~"high"))}

#PA_level definition
nhanes3<-nhanes3%>%mutate(PA_level = 
  case_when  
           ((PA <=150) ~ "low",
            (PA >150) & (PA <=1118) ~"intermediate",
            (PA >1118) ~"high"))

nhanes3<-nhanes3[!is.na(nhanes3$score)]
hanes3<-nhanes3[!is.na(nhanes3$risk)]
PA.c.protein<-nhanes3[!is.na(nhanes3$risk_level)]
```

## 1. Table 1 Baseline characteristics
```{r echo=FALSE, message=FALSE, warning=FALSE}


```

## 2. Scatterplot of the relationship between PA and C-protein
```{r echo=FALSE, message=FALSE, warning=FALSE}
PA.c.protein.plot<-ggplot(PA.c.protein,aes(x=PA,y=LBXCRP))+
  geom_point()
print(PA.c.protein.plot)
```

## 3. heatmap of PA and C-protein under different framingham score risks
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(PA.c.protein, aes(risk_level,PA, z = LBXCRP)) +
  stat_summary_2d()  +
  viridis::scale_fill_viridis()
```


\newpage

# Analysis

## Question 1: Is physical activity associated with the level of C-protein? If so, what is the exact relationship? Is it dose-response? 
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Split the data into training and test set
set.seed(123)
training.samples <- PA.c.protein$LBXCRP %>%
  createDataPartition(p = 0.8, list = FALSE)
train.data  <- PA.c.protein[training.samples, ]
test.data <- PA.c.protein[-training.samples, ]
# Build the model
model <- lm(LBXCRP ~ poly(PA, 8, raw = TRUE), data = train.data)
# Make predictions
predictions <- model %>% predict(test.data)
# Model performance
modelPerfomance<-data.frame(RMSE = RMSE(predictions, test.data$LBXCRP),
           R2 = R2(predictions, test.data$LBXCRP))

summary(lm(LBXCRP ~ poly(PA, 8, raw = TRUE), data = train.data))
print(modelPerfomance)

ggplot(train.data, aes(PA, LBXCRP) ) +
  geom_point() +
  stat_smooth(method = lm, formula = y ~ poly(x, 8, raw = TRUE))
```

```{r message=FALSE, warning=FALSE, include=FALSE}
mod_lm <- gam(LBXCRP ~ PA, data=PA.c.protein)
summary(mod_lm)
```

## Question 2: Are different levels of framingham scores the effect modifiers for this relationship between physical activity and C-protein? How do they modify the relationship?
```{r echo=FALSE, message=FALSE, warning=FALSE}
mod_gam3 <- gam(LBXCRP ~ te(PA, risk), data=PA.c.protein)
summary(mod_gam3)
vis.gam(mod_gam3, type='response', plot.type='persp',theta=30, n.grid=500, border=NA)
visreg2d(mod_gam3, xvar='risk', yvar='PA', scale='response')
```

\newpage

# Summary and Conclusions


\newpage

# References
<add references here if relevant, otherwise delete this section> 
